---
title: "Final Presentation"
author: "Author: Cameron Lyons"
date: "`r format(Sys.time(), '%B %d, %Y @ %I:%M %p')`"
output: ioslides_presentation
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE, # hide all the code
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```

## Packages Used

I used dplyr for data manipulation and plotting.

```{r, echo = TRUE}
library(dplyr)
```

I need readr to load in master file.

```{r, echo = TRUE}
library(readr)
```

I need ggplot to create spray chart and heat map.

```{r, echo = TRUE}
library(ggplot2)
```

I need scales to format the heat map.

```{r, echo = TRUE}
library(scales)
```

## Appalachian State Baseball Dataset

I compiled my dataset using this semester's TrackMan data collection with Appalachian State Baseball. The dataset was last updated November 6, 2022, for our last scrimmage of the Fall.

**Variables Used in the Dataset:**

* `PitcherTeam`: pitching team
* `BatterTeam`: batting team
* `Bearing`: axis of ball in play (dead center = 0)
* `Distance`: ball in play distance (feet)
* `PlayResult`: outcome of play
* `TaggedHitType`: hit type of ball in play  

## 

* `PlateLocSide`: plate location side (feet)
* `PlateLocHeight`: plate location height (feet)
* `ExitSpeed`: velocity of ball off contact (mph)
* `PitchCall`: what happened to the pitch

```{r}
master <- read_csv("2022-2023.csv", col_types = cols(Date = col_date(format = "%m/%d/%Y")))

master$Bearing <- ((master$Bearing * -1) + 90)*pi/180
master$x <- master$Distance*cos(master$Bearing)
master$y <- master$Distance*sin(master$Bearing)
bases <- data.frame(xa = c(0, 90/sqrt(2), 0, -90/sqrt(2), 0, rep(0)),
                    ya = c(0, 90/sqrt(2), 2 * 90/sqrt(2), 90/sqrt(2), 0, rep(0)))

batting <- master %>%
  filter(BatterTeam == "APP_MOU" | BatterTeam == "APP_PRA") %>%
  filter(!PlayResult == "Undefined" & !TaggedHitType == "Bunt")

x <- c(-.83,.83,.83,-.83,-.83)
y <- c(1.6,1.6,3.3,3.3,1.6)
sz <- data_frame(x, y)

x1 <- c(-.95, -.95, 0, .95, .95, -.95)
y1 <- c(0, .15, .25, .15, 0, 0)
hp <- data_frame(x1, y1)

x2 <- c(1.61, 3.11, 3.11, 1.61, 1.61)
y2 <- c(0, 0, 5.5, 5.5, 0)
rhb <- data_frame(x2, y2)

x3 <- c(-1.61, -3.11, -3.11, -1.61, -1.61)
y3 <- c(0, 0, 5.5, 5.5, 0)
lhb <- data_frame(x3, y3)

x4 <- c(0, 0, 1.08, 1.08, .83, 1.08, 1.08, 0, 0, 0, -1.08, -1.08, -.83, -1.08, -1.08, 0) 
y4 <- c(3.3, 3.55, 3.55, 2.497, 2.497, 2.497, 1.35, 1.35, 1.6, 1.35, 1.35, 2.497, 2.497, 2.497, 3.55, 3.55)
bllz <- data_frame(x4, y4)

x5 <- c(-.5, -.5, .5, .5, -.5)
y5 <- c(2, 3, 3, 2, 2)
innersix <- data_frame(x5, y5)

x6 <- c(-.95, .95, .95, -.95)
y6 <- c(2.233, 2.233, 2.866, 2.866)
szv <- data_frame(x6, y6)

x7 <- c(-0.316, -0.316, 0.316, 0.316)
y7 <- c(1.6, 3.5, 3.5, 1.6)
szh <- data_frame(x7, y7)

x8 <- c(0, 0, 1.267, 1.267, .95, 1.267, 1.267, 0, 0, 0, -1.267, -1.267, -.95, -1.267, -1.267, 0)
y8 <- c(3.5, 3.817, 3.817, 2.55, 2.55, 2.55, 1.283, 1.283, 1.6, 1.283, 1.283, 2.55, 2.55, 2.55, 3.817, 3.817)
bllzhm <- data_frame(x8, y8)

x9 <- c(-.95,.95,.95,-.95,-.95)
y9 <- c(1.6,1.6,3.5,3.5,1.6)
szhm <- data_frame(x9, y9)

 exits <- batting %>%
      filter(PitchCall == "InPlay" & ExitSpeed & !batting$TaggedHitType == "Bunt")
    
    tr <- exits %>%
      filter(PlateLocSide >= .316 & PlateLocSide <= .95 & PlateLocHeight >= 2.886 & PlateLocHeight <= 3.5) %>%
      summarise(mean = mean(ExitSpeed, na.rm = TRUE))
    t <- exits %>%
      filter(PlateLocSide <= .316 & PlateLocSide >= -.316 & PlateLocHeight >= 2.886 & PlateLocHeight <= 3.5) %>%
      summarise(mean = mean(ExitSpeed, na.rm = TRUE))
    tl <- exits %>%
      filter(PlateLocSide <= -.316 & PlateLocSide >= -.95 & PlateLocHeight >= 2.886 & PlateLocHeight <= 3.5) %>%
      summarise(mean = mean(ExitSpeed, na.rm = TRUE))
    mr <- exits %>%
      filter(PlateLocSide >= .316 & PlateLocSide <= .95 & PlateLocHeight <= 2.886 & PlateLocHeight >= 2.233) %>%
      summarise(mean = mean(ExitSpeed, na.rm = TRUE))
    m <- exits %>%
      filter(PlateLocSide <= .316 & PlateLocSide >= -.316 & PlateLocHeight <= 2.886 & PlateLocHeight >= 2.233) %>%
      summarise(mean = mean(ExitSpeed, na.rm = TRUE))
    ml <- exits %>%
      filter(PlateLocSide <= -.316 & PlateLocSide >= -.95 & PlateLocHeight <= 2.886 & PlateLocHeight >= 2.233) %>%
      summarise(mean = mean(ExitSpeed, na.rm = TRUE))
    br <- exits %>%
      filter(PlateLocSide >= .316 & PlateLocSide <= .95 & PlateLocHeight <= 2.223 & PlateLocHeight >= 1.6) %>%
      summarise(mean = mean(ExitSpeed, na.rm = TRUE))
    b <- exits %>%
      filter(PlateLocSide <= .316 & PlateLocSide >= -.316 & PlateLocHeight <= 2.223 & PlateLocHeight >= 1.6) %>%
      summarise(mean = mean(ExitSpeed, na.rm = TRUE))
    bl <- exits %>%
      filter(PlateLocSide <= -.316 & PlateLocSide >= -.95 & PlateLocHeight <= 2.223 & PlateLocHeight >= 1.6) %>%
      summarise(mean = mean(ExitSpeed, na.rm = TRUE))
    btr <- exits %>%
      filter(PlateLocSide >= .95 & PlateLocHeight >= 2.55 | PlateLocSide >= 0 & PlateLocHeight >= 3.5) %>%
      summarise(mean = mean(ExitSpeed, na.rm = TRUE))
    btl <- exits %>%
      filter(PlateLocSide <= -.95 & PlateLocHeight >= 2.55 | PlateLocSide <= 0 & PlateLocHeight >= 3.5) %>%
      summarise(mean = mean(ExitSpeed, na.rm = TRUE))
    bbr <- exits %>%
      filter(PlateLocSide >= .95 & PlateLocHeight <= 2.55 | PlateLocSide >= 0 & PlateLocHeight <= 1.6) %>%
      summarise(mean = mean(ExitSpeed, na.rm = TRUE))
    bbl <- exits %>%
      filter(PlateLocSide <= -.95 & PlateLocHeight <= 2.55 | PlateLocSide <= 0 & PlateLocHeight <= 1.6) %>%
      summarise(mean = mean(ExitSpeed, na.rm = TRUE))
    
    xtr <- c(.623)
    ytr <- c(3.193)
    trdf <- data.frame(x = xtr, y = ytr, z = tr)
    xt <- c(0)
    yt <- c(3.193)
    tdf <- data.frame(x = xt, y = yt, z = t)
    xtl <- c(-.623)
    ytl <- c(3.193)
    tldf <- data.frame(x = xtl, y = ytl, z = tl)
    xmr <- c(.623)
    ymr <- c(2.55)
    mrdf <- data.frame(x = xmr, y = ymr, z = mr)
    xm <- c(0)
    ym <- c(2.55)
    mdf <- data.frame(x = xm, y = ym, z = m)
    xml <- c(-.623)
    yml <- c(2.55)
    mldf <- data.frame(x = xml, y = yml, z = ml)
    xbr <- c(.623)
    ybr <- c(1.917)
    brdf <- data.frame(x = xbr, y = ybr, z = br)
    xb <- c(0)
    yb <- c(1.917)
    bdf <- data.frame(x = xb, y = yb, z = b)
    xbl <- c(-.623)
    ybl <- c(1.917)
    bldf <- data.frame(x = xbl, y = ybl, z = bl)
    hmsdf <- rbind(trdf, tdf, tldf, mrdf, mdf, mldf, brdf, bdf, bldf)
    hmsdf[is.na(hmsdf)] <- 0
    xbtr <- c(.634)
    ybtr <- c(3.184)
    btrdf <- data.frame(x = xbtr, y = ybtr, z = btr)
    xbtl <- c(-.634)
    ybtl <- c(3.184)
    btldf <- data.frame(x = xbtl, y = ybtl, z = btl)
    xbbr <- c(.634)
    ybbr <- c(1.917)
    bbrdf <- data.frame(x = xbbr, y = ybbr, z = bbr)
    xbbl <- c(-.634)
    ybbl <- c(1.917)
    bbldf <- data.frame(x = xbbl, y = ybbl, z = bbl)
    hmbdf <- rbind(btrdf, btldf, bbrdf, bbldf)
    hmbdf[is.na(hmbdf)] <- 0
```

##

```{r, fig.align = "center", fig.width = 7.5, fig}
ggplot(bases, aes(x = xa, y = ya)) +
      geom_path() +
      geom_point(data = batting, x = batting$x, y = batting$y,
                 aes(color = factor(batting$PlayResult))) +
      labs(color = "Play Result") +
      geom_segment(x = 0, xend = -233.35, y = 0, yend = 233.35) +
      geom_segment(x = -233.35, xend = -160.59, y = 233.35, yend = 340) +
      geom_segment(x = -160.59, xend = -49.36, y = 340, yend = 401.98) +
      geom_segment(x = -49.36, xend = 49.36, y = 401.98, yend = 401.98) +
      geom_segment(x = 49.36, xend = 157.67, y = 401.98, yend = 338.05) +
      geom_segment(x = 157.67, xend = 229.81, y = 338.05, yend = 229.81) +
      geom_segment(x = 229.81, xend = 0, y = 229.81, yend = 0) +
      xlim(-250, 250) + ylim(-10, 500) +
      theme(axis.line = element_blank(),
            axis.text.x = element_blank(),
            axis.text.y = element_blank(),
            axis.ticks = element_blank(),
            axis.title.x = element_blank(),
            axis.title.y = element_blank(),
            panel.background = element_blank(),
            panel.border = element_blank(),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            plot.background = element_blank()) +
      coord_equal() +
      scale_color_discrete(limits = c("Single" ,"Double", "Triple", "HomeRun", "Out", "FieldersChoice", "Error"))
```

This is a spray chart of every ball in play at Jim and Bettie Smith Stadium, Fall of 2022.

##

```{r, fig.align = "center", fig.width = 7.5}
ggplot(hmsdf, aes(x, y)) +
  geom_tile(data = hmbdf, aes(fill = mean), width = 1.267, height = 1.267, na.value = "white") +
  geom_tile(aes(fill = mean), width = .633, height = .633, na.value = "white") +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 85,
                       limits = c(75, 95), oob = squish) +
  labs(fill = "Exit Speed") +
  geom_path(data = szhm, aes(x = x9, y = y9)) +
  geom_path(data = szv, aes(x = x6, y = y6)) +
      geom_path(data = szh, aes(x = x7, y = y7)) +
      geom_path(data = bllzhm, aes(x = x8, y = y8)) +
      geom_text(mapping = aes(x = .623, y = 3.193), label = paste(sprintf("%.2f", tr)), size = 2.5) +
      geom_text(mapping = aes(x = 0, y = 3.193), label = paste(sprintf("%.2f", t)), size = 2.5) +
      geom_text(mapping = aes(x = -.623, y = 3.193), label = paste(sprintf("%.2f", tl)), size = 2.5) +
      geom_text(mapping = aes(x = .623, y = 2.55), label = paste(sprintf("%.2f", mr)), size = 2.5) +
      geom_text(mapping = aes(x = 0, y = 2.55), label = paste(sprintf("%.2f", m)), size = 2.5) +
      geom_text(mapping = aes(x = -.623, y = 2.55), label = paste(sprintf("%.2f", ml)), size = 2.5) +
      geom_text(mapping = aes(x = .623, y = 1.917), label = paste(sprintf("%.2f", br)), size = 2.5) +
      geom_text(mapping = aes(x = 0, y = 1.917), label = paste(sprintf("%.2f", b)), size = 2.5) +
      geom_text(mapping = aes(x = -.623, y = 1.917), label = paste(sprintf("%.2f", bl)), size = 2.5) +
      geom_text(mapping = aes(x = .95, y = 3.659), label = paste(sprintf("%.2f", btr)), size = 2.5) +
      geom_text(mapping = aes(x = -.95, y = 3.659), label = paste(sprintf("%.2f", btl)), size = 2.5) +
      geom_text(mapping = aes(x = .95, y = 1.442), label = paste(sprintf("%.2f", bbr)), size = 2.5) +
      geom_text(mapping = aes(x = -.95, y = 1.442), label = paste(sprintf("%.2f", bbl)), size = 2.5) +
      geom_path(data = rhb, aes(x = x2, y = y2), color = "lightgray") +
      geom_path(data = lhb, aes(x = x3, y = y3), color = "lightgray") +
      geom_text(mapping = aes(x = 2.36, y = 2.75), label = paste("RHB"), color = "lightgray") +
      geom_text(mapping = aes(x = -2.36, y = 2.75), label = paste("LHB"), color = "lightgray") +
      geom_polygon(hp, mapping = aes(x = x1, y = y1), color = "black", fill = "white") +
      coord_equal() +
      ggtitle("Exit Velocities Heat Map") +
      coord_cartesian(xlim = c(-4.5, 4.5), ylim = c(0, 5.5)) +
      theme(axis.line = element_blank(),
            axis.text.x = element_blank(),
            axis.text.y = element_blank(),
            axis.ticks = element_blank(),
            axis.title.x = element_blank(),
            axis.title.y = element_blank(),
            panel.background = element_blank(),
            panel.border = element_blank(),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            plot.background = element_blank(),
            legend.title = element_blank())
```

Exit Velocities in each zone. Inner 9 is strike zone. Outer 4 is a ball.